Quantifying the home ground advantage with `hyper2` and `hyper3` objects

### 2025-02-06 inst/home_advantage.Rmd
consistentify "time to process" information
home_away_table()

(takes about fifteen minutes to run without cache).

To cite the `hyper2` package in publications, please use
@hankin2017_rmd.  Here I analyse a small synthetic dataset for home
ground advantage, then go on to discuss a dataset presented by
Agresti.  But first, the smallest nontrivial example is a 2x2 case:

@@

>>>>

<font size="1"> (takes about fifteen minutes to run without cache)
</font>

To cite the `hyper2` package in publications, please use
@hankin2017_rmd.  Here I analyse a small synthetic dataset for home
ground advantage, then go on to discuss a dataset presented by
Agresti.  At the end, I show how a results table can be created from
scratch using bespoke function `home_away_table()`.

But first, the smallest nontrivial example is a 2x2 case:

@@

# Creating a home/away results table from scratch

Function `home_away_table()` takes a data frame of results and
produces a table amenable to analysis by `home_away()` or
`home_away3()`.

```{r showdata}
a <- read.table("celtic_rangers_livingston_falkirk.txt", header=TRUE)
head(a)
(aF <- home_away_table(a, give=FALSE))
(aT <- home_away_table(a, give=TRUE))
```

```{r doscottish,cache=TRUE}
(H <- home_away(aF))
(mH <- maxp(H))
```

```{r testrangersandceltic, cache=TRUE}
samep.test(H,c("Rangers","Celtic"))

```{r  testhomezero, cache=TRUE}
specificp.gt.test(H,"home",0)
```

```{r dohome3, cache=TRUE}
home_away3(aF,lambda=1.88)
maxp(home_away3(aF,lambda=1.88),give=1)
maxp(home_away3(aF,lambda=1.11),give=1)
```

```{r calcsuppscotty,cache=TRUE}
f <- function(l){ maxp(home_away3(aF,l),give=TRUE,n=1)$likes}
lam <- seq(from=1.4,to=2.3,len=10)
date()
loglike <- sapply(lam,f)
date()
```

```{r optscotty,cache=TRUE}
(o <- optimize(f,c(1.1,4.6),maximum=TRUE))
```

```{r scottychisq,cache=TRUE}
o
maxL <- f(o$maximum)
W <- maxL - f(1)    # f(1) being the likelihood of the null
W
pchisq(2*W,df=1,lower.tail=FALSE)
```

```{r scotmax}
(H <- home_away3(home_away_table(a,give=FALSE),o$maximum))  # o$maximum ~= 1.353
maxp(H,give=1)
```

```{r}
plot(lam,loglike-max(loglike),type='b')
abline(h=c(0,-2))
abline(v=1)
grid()
```

```{r rememberhyper2}
(mH2 <- round(mH*1000))
```

```{r sfasdf, cache=TRUE}
o$maximum
mH3 <- maxp(home_away3(home_away_table(a,give=FALSE),o$maximum),give=FALSE)
round(mH3*1000)
```

Now consider Celtic vs Falkirk, Celtic at home and also  Falkirk at home.

### (Celtic at home) {-}
* hyper2: P(Celtic;Falkirk) = $\frac{390+67; 56}{390 + 56 + 67}\simeq (0.89; 0.11)$
* hyper3: P(Celtic;Falkirk) = $\frac{1.77*412; 74}{1.77*412+74}\simeq(0.92; 0.08)$

actual scoreline:  61-8 or about $(0.88; 0.12)$

###  (Celtic away)
* hyper2: P(Celtic;Falkirk) = $\frac{390; 56+67}{390 + 56 + 67}\simeq (0.76;0.24)$
* hyper3: P(Celtic;Falkirk) = $\frac{412; 1.77*74}{412 + 1.77*74}\simeq(0.76; 24)$

actual scoreline: 43-17 or about $(0.72; 0.28)$
